\documentclass{liturgy}

\begin{document}

\begin{center}
  \huge
  A Commination

  \large
  or Denouncing of \textsc{God}'s Anger and Judgements against Sinners
\end{center}

\instruction{After morning prayer the leader shall say}
\begin{leader}
  In the early Church there was a godly discipline that those who had committed notorious sin should make open penance at the beginning of Lent that they may openly recommit themselves to the \textsc{Lord} and so their souls might be saved and that others may, by their example, share in their penitence.

  Instead however, it is thought to be good that at this time we should declare general sentences of \textsc{God}'s cursing of impenitent sinners, taken from Deuteronomy 27 and other passages of scripture, that we should be moved to earnest and true repentance; might be fortified against these snares of the enemy and grow ever more aware of our need for \textsc{God}'s mercy and forgiveness.
\end{leader}

\vgap

\instruction{Then the following passages from Deuteronomy 27 shall be read}
\begin{responses}
  < Cursed is anyone who makes an idol --- a thing detestable to the Lord, the work of skilled hands --- and sets it up in secret.
  > Amen.
  < Cursed is anyone who dishonours their father or mother.
  > Amen.
  < Cursed is anyone who moves their neighbour's boundary stone.
  > Amen.
  < Cursed is anyone who leads the blind astray on the road.
  > Amen.
  < Cursed is anyone who withholds justice from the foreigner, the fatherless or the widow.
  > Amen.
  < Cursed is anyone who sleeps with his neighbours' wife.
  > Amen.
  < Cursed is anyone who kills their neighbour secretly.
  > Amen.
  < Cursed is anyone who accepts a bribe to kill an innocent person.
  > Amen.
  < Cursed is anyone who does not uphold the words of this law by carrying them out.
  > Amen.\end{responses}

\vgap

\begin{leader}
  \noindent
  Now, keeping in mind those that the \textsc{Lord} has cursed and remembering the judgement which we deserve, let us turn to the \textsc{Lord} our God with contrite and meek hearts; in sorrow for and lamenting our sinful life, acknowledging and confessing our offences, and seeking to change our hearts and lives.
  For now the axe is against the tree, and every tree that does not bear good fruit will be cut down and cast into the fire.
  It is a fearful thing to fall into the hands of the living \textsc{God} for when he comes down from heaven to judge the wickedness of those on earth he will pour down rain and bring snares, fire and brimstone, storm and tempest upon the sinners.
  But who can withstand the day of judgement, who can stand before the \textsc{Lord}? He will gather up the wheat but the chaff he shall burn with an unquenchable fire.
  The day of the \textsc{Lord} will come like a thief in the night; when people say, Peace, and all is well then shall come upon them a sudden destruction as a woman is in pain during childbirth and no one shall escape it.
  Then, on the day of judgement, the wrath of the \textsc{Lord} shall be seen on those who, being obstinate sinners, have heaped it upon themselves through the stubbornness of their hearts, despising the goodness, patience and long-sufferance of \textsc{God} as he continually calls them to repentance.
  Then, says the \textsc{Lord}, they shall call upon me but I will not hear, they shall seek me but they will not find me for they hated knowledge and did not fear the \textsc{Lord} but rather abhorred my counsel and despised my correction.
  Then it shall be too late to knock for the door shall be shut and too late to cry for mercy for it shall be the time of justice.
  Then the terrible voice, which pronounces most just judgement, shall say to them "Go, you who are cursed, into the everlasting fires which have been prepared for the devil and his angels".
  Therefore, during this day of salvation, we must work, for the night is coming when no one can work.
  But, while we have the light, let us believe in the light and walk as children of the light that we may not be cast into the darkness where there is weeping and gnashing of teeth.
  Let us not abuse the goodness of \textsc{God}, who, in his mercy, calls us to amendment and who, in his endless pity, promises us forgiveness of all that which is past if we, with perfect and true hearts, turn back to him.
  For though our sins are red as scarlet they shall be made white as snow and though they be like purple they shall be make white as wool.
  Turn from your wickedness, says the \textsc{Lord}, and your sin shall not be your destruction.
  Let go of all ungodliness in your lives; make a new heart and a new spirit within yourselves that you may be saved.
  For I take no pleasure in any death so turn then and you shall live.
  For although we have sinned we have an Advocate with the Father, Jesus Christ the righteous; and he is the propitiation for our sins.
  For he was wounded for our offences and suffered for our wickedness.
  Therefore let us return to him, who receives all true penitent sinners, assuring ourselves that he is ready to receive us and most willing to pardon us, if we come to him with faithful repentance, submit ourselves to him and from henceforth walk in his ways; if we take his easy yoke and light burden upon us to follow him in lowliness, patience and charity, and be ordered by the governance of his Holy Spirit; seeking always his glory, and serving him duly in our vocation with thanksgiving.
  If we do this then Christ will deliver us from the curse of the law, and from the judgement due those who are on his left hand and he shall set us on his right hand, giving us the gracious blessing of his Father, and commanding us to take possession of his glorious kingdom into which he promises, though his infinite mercy, to bring us. \inlineall{Amen.}
\end{leader}

\instruction{Then, kneeling, Psalm 51 shall be read}
\begin{canticle}
  Have mercy on me, O God, in your great goodness; &
  according to the abundance of your compassion \\
  blot out my offences.

  Wash me thoroughly from my wickedness  &
  and cleanse me from my sin.

  For I acknowledge my faults  &
  and my sin is ever before me.

  Against you only have I sinned  &
  and done what is evil in your sight,

  So that you are justified in your sentence  &
  and righteous in your judgement.

  I have been wicked even from my birth,  &
  a sinner when my mother conceived me.

  Behold, you desire truth deep within me  &
  and shall make me understand wisdom \\
  in the depths of my heart.

  Purge me with hyssop and I shall be clean;  &
  wash me and I shall be whiter than snow.

  Make me hear of joy and gladness,  &
  that the bones you have broken may rejoice.

  Turn your face from my sins  &
  and blot out all my misdeeds.

  Make me a clean heart, O God,  &
  and renew a right spirit within me.

  Cast me not away from your presence  &
  and take not your holy spirit from me.

  Give me again the joy of your salvation  &
  and sustain me with your gracious spirit;

  Then shall I teach your ways to the wicked  &
  and sinners shall return to you.

  Deliver me from my guilt, O God, \\
  the God of my salvation,  &
  and my tongue shall sing of your righteousness.

  O Lord, open my lips  &
  and my mouth shall proclaim your praise.

  For you desire no sacrifice, else I would give it;  &
  you take no delight in burnt offerings.

  The sacrifice of God is a broken spirit;  &
  a broken and contrite heart, O God, you will not despise.

  O be favourable and gracious to Zion;  &
  build up the walls of Jerusalem.

  Then you will accept sacrifices offered in righteousness, \\
  the burnt offerings and oblations;  &
  then shall they offer up bulls on your altar.
\end{canticle}

\vgap

\begin{responses}
  > Glory to the Father and to the Son
  > and to the Holy Spirit;
  > as it was in the beginning is now
  > and shall be for ever. Amen.
  -
  < Lord, have mercy upon us.
  > Christ, have mercy upon us.
  < Lord, have mercy upon us.
  -
  > Our Father in heaven,
  > hallowed be your name,
  > your kingdom come,
  > your will be done,
  > on earth as in heaven.
  > Give us today our daily bread.
  > Forgive us our sins
  > as we forgive those who sin against us.
  > Lead us not into temptation
  > but deliver us from evil.
  > For the kingdom, the power,
  > and the glory are yours
  > now and for ever.
  > Amen.
  | (or)
  > Our Father, who art in heaven,
  > hallowed be thy name;
  > thy kingdom come;
  > thy will be done;
  > on earth as it is in heaven.
  > Give us this day our daily bread.
  > And forgive us our trespasses,
  > as we forgive those who trespass against us.
  > And lead us not into temptation;
  > but deliver us from evil.
  > For thine is the kingdom,
  > the power, and the glory
  > for ever and ever.
  > Amen.
  -
% < O \textsc{Lord}, save your servants;
  < O Lord, save your servants;
  > Who put their trust in you.
  < Send them help from above.
  > And mightily defend them for evermore.
% < Help us, O \textsc{God} out Saviour.
  < Help us, O God out Saviour.
  > And deliver us for the glory of your Name; be merciful to us sinners for your Name's sake.
  < O Lord, hear our prayer.
  > And hear our cry for mercy.
  < Let us pray
\end{responses}

\begin{leader}
  O \textsc{Lord}, mercifully hear our prayers, and spare all those who confess their sins to you; that as their consciences are accused by sin they may be absolved by your merciful pardon; through Christ our \textsc{Lord}. \inlineall{Amen.}

  O most mighty \textsc{God}, and merciful Father, who had compassion on all men, and hates nothing that you have made, who doesn't desire the death of a sinner, but that he should turn from his sin, and be saved: Mercifully forgive us our trespasses and receive and comfort us who are grieved and wearied with the burden of our sins. Your property is always to have mercy and you alone has the power to forgive sins. Spare us therefore, good \textsc{Lord}, spare your people, whom you have redeemed, and refrain from judgement of your servants, who are but vile dust, miserable sinners; but turn your anger from us, who meekly acknowledge our vileness, and truly repent of all faults, and so make haste to help us in this world, that we may evermore live with you in the world to come; through Jesus Christ our \textsc{Lord}. \inlineall{Amen.}
\end{leader}
\begin{all}
  Restore us to yourself, \textsc{Lord}, that we may turn to you. Be favourable, \textsc{Lord}, be favourable to us as we turn to you in weeping, fasting and praying. For you are a merciful \textsc{God}, full of compassion, long-suffering, and of great pity. You spare us when we deserve punishment and in your wrath you show mercy. Spare your people, O \textsc{Lord}, spare them and do not let your heritage be brought to confusion. Hear us, O \textsc{Lord}, for great is your mercy, and in the multitude of your mercies look upon us; through the merits and mediation of your blessed Son, Jesus Christ our \textsc{Lord}. Amen.
\end{all}
\begin{leader}
  The Lord bless us and keep us; the Lord make his face shine upon us and give us peace, now and for evermore. \inlineall{Amen.}
\end{leader}

\vfill

{
  \tiny\noindent
  Elements taken from Common Worship: Services and Prayers for the Church of England (2000)
}

\end{document}

